Behavior Driven Development in Haskell

### 2012-05-29 paper.tex
More work on "Problem"

Many Haskell programmers tend not to write test code even though the
strong type system of Haskell cannot ensure correct semantics.
Without test code, it is difficult for users to understand usage of
functions and it is difficult for developers to maintain semantics
while extending software. To enhance the value of test code, we
propose to carry out behavior driven development with two tools:
doctest integrated with QuickCheck for user documentation and Hspec
with combinators extended for developer documentation. With our
method, specification can be used for documentation and automatic
testing. This kill-three-birds-with-one-stone aspect would motivate
Haskellers to write test code.

@@

\begin{quote}\begin{itemize}
    \item
        BDD tools for other programming languages use code examples as
        evidence for behavior.
    \item
        When we apply BDD to Haskell we want to use properties and
        examples to provide evidence for behavior.
    \item
        Hspec already has support for both.
    \item
        Haddock has markup support for interactive examples, and
        Doctest can verify them.
    \item
        But it is not possible to use properties in Haddock comments
        and automatically test them.
\end{itemize} \end{quote}

\begin{quote}
Hspec can be used to document the behavior of a system in a way that
is similar to a specification.  All facets of the system should be
described with examples and properties, including every possible
corner case.

We think that some evidence is useful as user documentation, hence we
want to be able to include it in Haddock comments.  But there is also
evidence that is not useful as user documentation.  E.g. that a
particular \texttt{Monoid} instance is a monoid in the mathematical
sense is self-evident.  So we do not want to include any examples or
behaviors for that into user documentation, but we still want to be
able to automatically test it.

We propose to use Doctest to test evidence that is useful as user
documentation, and Hspec for evidence that is not useful as user
documentation.
\end{quote}

BDD tools should be treat both of them and provide functionality
to extract evidence for users as user documentation.
If we can archive this, BDD tools can be used for
triple purpose: design, documentation and automatic testing.

@@

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Open problems} %% Future work?

>>>>

Many Haskell programmers tend not to write test code even though the
strong type system of Haskell cannot always ensure correct semantics.
Without test code, it is difficult for users to understand usage of
functions and it is difficult for developers to maintain semantics
while extending software. To enhance the value of test code, we
propose to carry out behavior driven development with two tools:
doctest integrated with QuickCheck for user documentation and Hspec
with combinators extended for developer documentation. With our
method, specification can be used for documentation and automatic
testing. This kill-three-birds-with-one-stone aspect would motivate
Haskellers to write test code.

@@

Hspec can be used to document the behavior of a system in a way that
is similar to a specification.  All facets of the system should be
described with examples and properties, including every possible
corner case.

We think that some evidence is useful as user documentation, hence we
want to be able to include it in Haddock comments.  But there is also
evidence that is not useful as user documentation.  E.g. that a
particular \texttt{Monoid} instance is a monoid in the mathematical
sense is self-evident.  So we do not want to include any evidence for
that into user documentation, but we still want to automatically test
it.

We propose to use Doctest and Haddock to test evidence that is useful
as user documentation, and Hspec for evidence that is not useful as
user documentation.

\begin{quote}
BDD tools should be treat both of them and provide functionality
to extract evidence for users as user documentation.
If we can archive this, BDD tools can be used for
triple purpose: design, documentation and automatic testing.
\end{quote}

@@

\begin{quote}\begin{itemize}
    \item
        Haddock has markup support for interactive examples, and
        Doctest can verify them.
    \item
        But it is not possible to use properties in Haddock comments
        and automatically test them.
\end{itemize} \end{quote}

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Open problems and future work}
